---
title: "Untitled"
author: "Andrew W. Park"
date: '2022-10-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Required packages
```{r loadLibs}
library(dplyr)
library(maptools)
library(raster)
library(metacom)
```


```{r loadData}
load("get_db.Rda")

#check for UTM typos
Vial.ID.issues <- db$Vial.ID[which(db$UTM.1==db$UTM.2)]

#fix a few database typos confirmed by email from Stacey Vigil to Andrew Park (11/11/2022)
#AP's copy has UTM.2=UTM.1, Stacey looked up correct UTM.2 in original data sheets
db$UTM.2[which(db$Vial.ID=="FL0276")] <- 2957038
db$UTM.2[which(db$Vial.ID=="FL0315")] <- 2957057
db$UTM.2[which(db$Vial.ID=="FL0567")] <- 3001295
db$UTM.2[which(db$Vial.ID=="FL0568")] <- 3002091
db$UTM.2[which(db$Vial.ID=="FL0816")] <- 3034831
db$UTM.2[which(db$Vial.ID=="FL0817")] <- 3034831
db$UTM.2[which(db$Vial.ID=="FL0818")] <- 3034831


#here I [AP] just convert UTM lat/long using http://rcn.montana.edu/resources/Converter.aspx
#FL is un UTM zone 17 (northern hemisphere)
#vial FL0276
db$Lat[which(db$Vial.ID=="FL0276")] <- 26.73170
db$Long[which(db$Vial.ID=="FL0276")] <- -81.90239
#vial FL0315
db$Lat[which(db$Vial.ID=="FL0315")] <- 26.73190
db$Long[which(db$Vial.ID=="FL0315")] <- -81.89777
#vial FL0567
db$Lat[which(db$Vial.ID=="FL0567")] <- 27.12999
db$Long[which(db$Vial.ID=="FL0567")] <- -82.08142
#vial FL0568
db$Lat[which(db$Vial.ID=="FL0568")] <- 27.13710
db$Long[which(db$Vial.ID=="FL0568")] <- -82.09118
#vial FL0816
db$Lat[which(db$Vial.ID=="FL0816")] <- 27.43069
db$Long[which(db$Vial.ID=="FL0816")] <- -82.31852
#vial FL0817
db$Lat[which(db$Vial.ID=="FL0817")] <- 27.43069
db$Long[which(db$Vial.ID=="FL0817")] <- -82.31852
#vial FL0818
db$Lat[which(db$Vial.ID=="FL0818")] <- 27.43069
db$Long[which(db$Vial.ID=="FL0818")] <- -82.31852

db %<>% dplyr::filter(Lat<35) # four sites are up in Michigan and spatially disconnected from this metacommunity analysis

```






## Creating Centroids for WMAs
```{r wmaLatLong}

centroids <- db |> 
  group_by(Site) |> 
  summarize(cLat=mean(Lat),cLon=mean(Long))


save(centroids,file="get_siteCentroids.Rda")

```





```{r}

sites <- unique(db$Site)
sub.db <- db[,11:62]

#for loop that outputs interaction matrix combining all years
ret <- matrix(0, ncol= 52, nrow=length(sites)) 
latitude=vector(); longitude=vector()
for(i in 1:length(sites)){
  temp=sub.db[which(db$Site==sites[i]),]	
  ret[i,] = colSums(temp)
  latitude[i] = mean(db[which(db$Site == sites[i]),'Lat'])
  longitude[i] = mean(db[which(db$Site == sites[i]),'Long'])
}
rownames(ret)=sites; colnames(ret)=colnames(sub.db)

#ret.lat=ret[order(latitude),]
ret.nozero <- ret[-which(rowSums(ret) == 0),]
ret.nozero <- ret.nozero > 0 
sites.nozero <- sites[-which(rowSums(ret)==0)]
latitude <- latitude[-which(rowSums(ret)==0)]
longitude <- longitude[-which(rowSums(ret)==0)]

```





### Default ordering based on RA 

```{r}
#pull out site scores from reciprocal averaging
scores <- OrderMatrix(ret.nozero, scores=1, outputScores=TRUE)

#Coherence 
coh.ret.nozero <- Coherence(ret.nozero > 0, order=TRUE, orderNulls=TRUE)
#Species Turnover
turn.ret.nozero <- Turnover(ret.nozero > 0, order=TRUE, orderNulls=TRUE)
#Boundary clumping
clump.ret.nozero <- BoundaryClump(ret.nozero > 0, order=TRUE)

```







### Spatial ordering

```{r}
ret.spatial <- ret.nozero[order(latitude),order(scores$speciesscores)]

#Coherence 
coh.ret.spatial <- Coherence(ret.spatial > 0, order=FALSE, sims=100, orderNulls=TRUE)
#Species Turnover
turn.ret.spatial <- Turnover(ret.spatial > 0, order=FALSE, orderNulls=TRUE)
#Boundary clumping
clump.ret.spatial <- BoundaryClump(ret.spatial > 0, order=FALSE)

```




Modularity of network

```{r}

getModularity <- function(x, sims, onlyZ=TRUE, useProjection=FALSE){
  require(metacom)
  require(igraph)
  nulls <- metacom::NullMaker(x, sims=sims, method='swap')
  iNulls <- lapply(nulls, function(x){
    igraph::graph_from_incidence_matrix(x)
  })
  ix <- igraph::graph_from_incidence_matrix(x)
  if(useProjection){
    ix <- igraph::bipartite_projection(ix)[[1]]
    iNulls <- lapply(iNulls, function(x){igraph::bipartite_projection(x)[[1]]})
  }
  Qnull <- sapply(iNulls, function(a){
    igraph::modularity(cluster_walktrap(a))
  })
  q <- igraph::modularity(cluster_walktrap(ix))
  zMod <- (q- mean(Qnull)) / (sd(Qnull))
  pMod <- 2 * pnorm(-abs(zMod))
  if(onlyZ){
    return(zMod)
  }else{
    return(c(Q=q, z=zMod, p=pMod))
  }
}

```








First pass modularity of the bipartite graph 

```{r}

mod1 <- getModularity(ret.spatial>0, sims=1000, onlyZ=FALSE)

```
     Q             z             p 
 3.182671e-01 4.545019e+00  5.493028e-06


```{r}

g <- igraph::graph_from_incidence_matrix(ret.spatial>0)
modules <- igraph::cluster_walktrap(g)

```




```{r}
#' @title Plot a graph object
#'
#' @param g graph object
#' @param colz colors to use
#' @param nodeSize Size of nodes
#' @param colVec What vertex attribute should color nodes? (default = NULL)
#' @param edgeWidth edge weights e.g., E(g)$weight? (default=1)
#' 
#' @return a graph plot

plotGraph <- function(g, lay=layout_nicely(g), 
	colz=1:5, nodeSize=10, colVec=NULL, edgeWidth=1){
	if(is.null(colVec)){
		colVec <- rep(1, length(V(g)))
	}
	plot(g, layout=lay, edge.width=edgeWidth, 
		vertex.size=nodeSize, directed=FALSE,
		vertex.color=colz[colVec], vertex.label=NA)
}

```



Visualize the network, calculating modularity on the bipartite network. I don't think this is _acutally_ what we did in the original manuscript, but I'm not sure. We're now finding 4 modules, not the 2 clumps we did before. However, if we project the bipartite matrix into a graph of sites connected by the number of shared vector species, we recover two modules just like before. 

```{r}

pdf('modularNetwork.pdf', height=4, width=4)
plotGraph(g, colz=viridis::viridis(4), 
  colVec=modules$membership)
dev.off()

```




If we simply project the bipartite graph into a site graph where connections are shared vectors. This recovers the two group results we saw previously.
ecology
```{r}

g2 <- bipartite_projection(g)[[1]]
#save(g2,file="get_g2.Rda")
mod2 <- getModularity(ret.spatial>0, sims=1000, onlyZ=FALSE, useProjection=TRUE)


modules2 <- cluster_walktrap(g2)

pdf('modularNetwork2.pdf', height=7, width=7)
plotGraph(g2, colz=viridis::viridis(4), 
  colVec=modules2$membership, edgeWidth=E(g2)$weight)
dev.off()

```


    Q             z             p 
 2.206874e-01 4.247183e+00  2.164752e-05










Make an image plot like the one currently in the manuscript, rows ordered by disease severity, modules identified with boxes. 

```{r}
# have not done. Don't see where hdScore is, so I'm holding off for now. 
colnames(ret.spatial)
rownames(ret.spatial)

```

